---
title: "p8105_hw3_slb2240"
author: "Samantha Brown"
date: "10/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
```

## Problem 1

Import and clean data:
```{r}
install_github("p8105/p8105.datasets")
library(p8105.datasets)
data(brfss_smart2010)

brfss_smart2010 = janitor::clean_names(brfss_smart2010) %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = as.factor(response))
```

**In 2002, which states were observed at 7 locations?**

```{r}
brfss_2002 = filter(brfss_smart2010, year == "2002") %>% 
  distinct(locationabbr, locationdesc) %>% 
  count(locationabbr) %>% 
  filter(n == 7)

brfss_2002
```

Connecticut, Florida, and North Carolina were observed at 7 locations.

**Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.**

geom_line(), with  group_by and summarize() commands for my plot. 


**Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.**

```{r}
brfss_NY = filter(brfss_smart2010, locationabbr == "NY") %>% 
  spread(key = response, value = data_value)

brfss_NY_excellent_prop = janitor::clean_names(brfss_NY) %>% 
  mutate(prop_excellent = excellent/sample_size) %>%
  group_by(year) %>% 
  summarize(mean_excellent = round(mean(prop_excellent, na.rm = TRUE), digits = 3),
            sd_excellent = round(sd(prop_excellent, na.rm = TRUE), digits = 3)) %>% 
  filter(year == c("2002", "2006", "2010"))

brfss_NY_excellent_prop

```

**For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.**

```{r}
brfss_location_state_prop = brfss_smart2010 %>% 
  spread(key = response, value = data_value) %>%
  janitor::clean_names() %>%
  group_by(year, locationabbr) %>%
  summarize(prop_excellent = round(mean(excellent/sample_size, na.rm = TRUE), digits = 2),
            prop_very_good = round(mean(very_good/sample_size, na.rm = TRUE), digits = 2),
            prop_fair = round(mean(fair/sample_size, na.rm = TRUE), digits = 2),
            prop_good = round(mean(good/sample_size, na.rm = TRUE), digits = 2),
            prop_poor = round(mean(poor/sample_size, na.rm = TRUE), digits = 2)) %>% 
  rename(state = locationabbr) %>% 
  gather(key = "response", value = "proportion_response", prop_excellent:prop_poor)
  
  
ggplot(brfss_location_state_prop, aes(x = state, y = proportion_response, color = year)) +
  geom_point(alpha = 0.2) +
  facet_grid(~response) +
  scale_color_viridis()
```
? How are we supposed to organize plots?


## Problem 2

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

```{r}
data(instacart)
```

The `instacart` dataset is comprised of anonymous 2017 online grocery orders with `r nrow(instacart)` observations of `r ncol(instacart)` variables. Because the dataset does not draw from neither randmomized users nor purchases, the results may not accurately represent a more generalized population of Instacart users. However, for the purpose of this homework, I will analyze trends in the `instacart` dataset.

I am curious about the following questions:
* Which department is most frequently ordered from? Which aisle?
* What is the average days since prior order?
* What proportion of Instacart users reordered products?

The snacks department is most frequently ordered from, while the alcohol deparment is ordered from the least. Meanwhile, the yogurt aisle is most frequently ordered from, while the air fresheners/candles aisle is ordered from the least.
```{r}
mode_dept = as.vector(instacart$department)
max(mode_dept)
min(mode_dept)

mode_aisle = as.vector(instacart$aisle)
max(mode_aisle)
min(mode_aisle)
```

On average, Instacart users tend to make orders every 15 days. The mean `days_since_prior_order` value is approximately 17 days, indicating that the number of days between orders is slightly positively skewed.
```{r}
median(instacart$days_since_prior_order)
mean(instacart$days_since_prior_order)
```

About `r round(sum(instacart$reordered)/nrow(instacart)*100, digits = 2)`% of customers reordered products.

## Describe some key variables, and give illstrative examples of observations.

**How many aisles are there, and which aisles are the most items ordered from?**

There are 134 aisles. The aisles in which the most items are ordered from are `fresh fruits`, `fresh vegetables`, `packaged cheese`, `packaged` `vegetables fuits`, and `yogurt`.

```{r}
n_aisles = instacart %>% 
  distinct(aisle) %>% 
  nrow()

most_freq_aisle = instacart %>%
  select(aisle, add_to_cart_order) %>% 
  group_by(aisle) %>% 
  summarize(sum(add_to_cart_order)) %>%
  top_n(5)
```

**Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.**

## Finish

**Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”**

The table below shows the top 5 most popular items in aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits,” by the sum of items added to cart.

```{r}
popular_items = instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  select(aisle, product_name, add_to_cart_order) %>% 
  group_by(aisle, product_name) %>% 
  summarize(sum(add_to_cart_order)) %>% 
  top_n(1)

popular_items
```

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r}
apples_coffee_dow = instacart %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>%
  summarize(round(mean(order_hour_of_day), digits = 0)) 

clean_apples_coffee = apples_coffee_dow %>% 
  rename(average_hour = "round(mean(order_hour_of_day), digits = 0)", day_of_week = order_dow) %>% 
  spread(key = day_of_week, value = average_hour) %>% 
  View()
```

## How do you get this as days of week? And times?

The table indicates that Instacart users tend to order Pink Lady apples around noon. Meanwhile, users generally order coffee ice cream in the afteroon. This makes sense because people likely think more about dessert in the afternoon.

## Problem 3

Load in dataset:
```{r}
data(ny_noaa)

ny_noaa = ny_noaa %>% 
  rename(snow_depth = snwd, temp_max = tmax, temp_min = tmin)
```

**Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue.**


The `ny_noaa` dataset has `r nrow(ny_noaa)` observations of `r ncol(ny_noaa)` variables. Each observation represents a New York weather station's recording of weather variables including the date in which the recording is made, precipitation (in tenths of mm), snowfall (in mm), depth of snow (in mm), maximum temperature (in tenths of degrees C), and minimum temperature (in tenths of degrees C). The dataset contains missing data because each station may only record a select few of the variables. I will explore the extent to which data is missing below:

* `r round(sum(is.na(ny_noaa$prcp))/nrow(ny_noaa), digits = 4)*100`% of precipitation data is missing.
* `r round(sum(is.na(ny_noaa$snow))/nrow(ny_noaa), digits = 4)*100`% of snowfall data is missing.
* `r round(sum(is.na(ny_noaa$snow_depth))/nrow(ny_noaa), digits = 4)*100`% of snow depth data is missing.
* `r round(sum(is.na(ny_noaa$temp_max))/nrow(ny_noaa), digits = 4)*100`% of maximum temperature data is missing.
* `r round(sum(is.na(ny_noaa$temp_min))/nrow(ny_noaa), digits = 4)*100`% of minimum temperature data is missing.

These numbers show that an extensive proportion of the data is missing. However, for the purpose of this homework I will continue working with the dataset.

```{r}
summary(ny_noaa$prcp)

summary(ny_noaa$snow)

summary(ny_noaa$snow_depth)

summary(as.numeric(ny_noaa$temp_max))

summary(as.numeric(ny_noaa$temp_min))
```


^^ Comment on summary statistics



**Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?**


**Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?**


**Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.**

