---
title: "p8105_hw3_slb2240"
author: "Samantha Brown"
date: "10/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cellranger)
library(devtools)
```

## Problem 1

Import and clean data:
```{r}
install_github("p8105/p8105.datasets")
library(p8105.datasets)
data(brfss_smart2010)

brfss_smart2010 = janitor::clean_names(brfss_smart2010) %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = as.factor(response))
```

**In 2002, which states were observed at 7 locations?**

```{r}
brfss_2002 = filter(brfss_smart2010, year == "2002") %>% 
  distinct(locationabbr, locationdesc) %>% 
  count(locationabbr) %>% 
  filter(n == 7)

brfss_2002
```

Connecticut, Florida, and North Carolina were observed at 7 locations.

**Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.**

geom_line(), with  group_by and summarize() commands for my plot. 


**Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.**

```{r}
brfss_NY = filter(brfss_smart2010, locationabbr == "NY") %>% 
  spread(key = response, value = data_value)

brfss_NY_excellent_prop = janitor::clean_names(brfss_NY) %>% 
  mutate(prop_excellent = excellent/sample_size) %>%
  group_by(year) %>% 
  summarize(mean_excellent = round(mean(prop_excellent, na.rm = TRUE), digits = 3),
            sd_excellent = round(sd(prop_excellent, na.rm = TRUE), digits = 3)) %>% 
  filter(year == c("2002", "2006", "2010"))

brfss_NY_excellent_prop

```

**For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.**

```{r}
brfss_location_state_prop = brfss_smart2010 %>% 
  spread(key = response, value = data_value) %>%
  janitor::clean_names() %>%
  group_by(year, locationabbr) %>%
  summarize(prop_excellent = round(mean(excellent/sample_size, na.rm = TRUE), digits = 2),
            prop_very_good = round(mean(very_good/sample_size, na.rm = TRUE), digits = 2),
            prop_fair = round(mean(fair/sample_size, na.rm = TRUE), digits = 2),
            prop_good = round(mean(good/sample_size, na.rm = TRUE), digits = 2),
            prop_poor = round(mean(poor/sample_size, na.rm = TRUE), digits = 2)) %>% 
  rename(state = locationabbr) %>% 
  gather(key = "response", value = "proportion_response", prop_excellent:prop_poor)
  
  
ggplot(brfss_location_state_prop, aes(x = state, y = proportion_response, color = year)) +
  geom_point(alpha = 0.2) +
  facet_grid(~response) +
  scale_color_viridis()

```
? How are we supposed to organize plots?


## Problem 2

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.





